State of the Practice for Medical Imaging Software Based on Open Source Repositories

### 2024-10-29 StateOfPractice/Papers/SOP_MI_CritReviewsBioMedEng/SOP_MI_OSR.tex
Revised abstract, now for full paper

%\author{Anonymous}
\author[CAS]{Spencer Smith}
\author[CAS]{Ao Dong}
\author[CAS]{Jacques Carette}
\author[EE]{Michael Noseworthy}

@@

\affiliation[ME]{organization={McMaster University, Electrical and Computer Engineering}, %Department and Organization
            addressline={1280 Main Street West}, 
            city={Hamilton},
            postcode={L8S 4K1}, 
            state={Ontario},
            country={Canada}}

@@

    We present the state of the practice for Medical Imaging (MI) software based
on data available in open source repositories. We selected 29 projects from 48
candidates and assessed 9 software qualities (installability, correctness/
verifiability, reliability, robustness, usability, maintainability, reusability,
understandability, and visibility/transparency) by answering 108 questions for
each. Using the Analytic Hierarchy Process (AHP) on the quantitative data, we
ranked the MI software.  The top five are \textit{3D Slicer}, \textit{ImageJ},
\textit{Fiji}, \textit{OHIF Viewer}, and \textit{ParaView}.  This is consistent
with the community's view, with four of these also appearing in the top five
using GitHub metrics (stars-per-year).  
In general, the quality and quantity of documentation present in a project
correlates quite well with its popularity.

>>>>

%\author{Anonymous}
\author[CAS]{Spencer Smith}
\author[CAS]{Ao Dong}
\author[CAS]{Jacques Carette}
\author[ECE]{Michael Noseworthy}

@@

\affiliation[ECE]{organization={McMaster University, Electrical and Computer Engineering}, %Department and Organization
            addressline={1280 Main Street West}, 
            city={Hamilton},
            postcode={L8S 4K1}, 
            state={Ontario},
            country={Canada}}

@@

We present the state of the practice for Medical Imaging (MI) software based on
data available in open source repositories. We selected 29 projects from 48
candidates and assessed 9 software qualities (installability, correctness/
verifiability, reliability, robustness, usability, maintainability, reusability,
understandability, and visibility/transparency) by answering 108 questions for
each. Using the Analytic Hierarchy Process (AHP) on the quantitative data, we
ranked the MI software.  The top five are \textit{3D Slicer}, \textit{ImageJ},
\textit{Fiji}, \textit{OHIF Viewer}, and \textit{ParaView}.  This is consistent
with the community's view, with four of these also appearing in the top five
using GitHub metrics (stars-per-year).  The quality and quantity of
documentation present in a project correlates quite well with its popularity.
Generally, MI software is in a healthy state as shown by the following: in the
repositories we observed 88\% of the documentation artifacts recommended by
research software development guidelines, 100\% of MI projects use version
control tools, and developers appear to use the common quasi-agile research
software development process. However, the current state of the practice
deviates from the existing guidelines because of the rarity of some recommended
artifacts (like a test plan, requirements specification, code of conduct, and
code style guidelines), low usage of continuous integration (17\% of the
projects), low use of unit testing (about 50\% of projects), and room for
improvement with documentation. From interviewing the developers, we identified
6 concerns: lack of development time, lack of funding, technology hurdles,
correctness, usability, maintainability, and reproducibility. Recommendations
for improving the state of the practice include the following: increase
documentation, increase testing by enriching datasets, increase continuous
integration usage, move to web applications, employ linters, use peer reviews,
and design for change.
